---
title: "Project2"
author: "Yiliang Yuan & Yiyu Lin"
date: "2022-11-26"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

***
## 1 Data Collection and Exploration
## (a)
Global climate change is an important scientific and public topic. In this topic, there’s a prediction that, in Arctic area, surface air temperatures have the strongest dependency on increasing atmospheric carbon dioxide levels. In further study of these dependency, the key is to detect the cloud condition. This is challenging problem if we apply two previous MISR operational algorithms since the similarity between remote sensing image of clouds and that of ice- or snow-covered surface. Thus, based on the new idea: searching for cloud-free surface, the goal of this study is to develop a computationally efficient algorithm with the combination of classification and clustering methods, but without the requiring human intervention
The data in this study represents 10 MISR orbits from Path 36, which is 360 km wide, and starting at Arctic until Antarctica. The time between each orbit is 16 days. In this path, there’re 180 blocks to collect data, but for each orbit case, only 6 blocks’ data will be taken. In these total 60 blocks, author removed 3 blocks whose cloud condition can be detected by previous MISR algorithm. Therefore, there’re total 57 blocks data unit with 7,114,248 1.1-km resolution pixels with 36 radiation measurements for each pixel in this study. Author hand-labeled these data, with “1” for cloudy, “-1” for not cloudy, and “0” for ambiguous.
There’re also three key features used in this study to differentiate surface pixels from cloudy pixels. The first is CORR. It is used to differentiate MISR image of the same scene from different MISR viewing directions. The second is SDAn, the standard deviation of MISR nadir camera pixel values across a scene. The last one, NDAI, is a normalized difference angular index. It can be used characterizes the changes in a scene with changes in the MISR view direction.
In conclusion, both new-developed ELCM and ELCM-QDA algorithm return a result with better accuracy and better coverage, when compared to the results generated from previous MISR algorithm. Moreover, ELCM-QDA algorithm can generate more information and perform consistently on probability prediction. This impact of this work is significant. Statisticians will play a more important role to in data analyses and statistical thinking will take more responsibility in solving modern scientific problems.


## (b)
```{r}
image1=read.csv('imagem1.txt',sep='',header=FALSE)
image2=read.csv('imagem2.txt', sep='',header=FALSE)
image3=read.csv('imagem3.txt', sep='',header=FALSE)
image <- rbind(image1,image2,image3)
colnames(image) = c('y','x','expert_label','NDAI','SD','CORR','DF','CF','BF','AF','AN')
```

```{r}
image1b=image[,c(1,2,3)]
image1b %>% group_by(expert_label) %>% summarise(per=n()/nrow(.)*100)
```
The percentage of pixels in cloud (1) class is 36.77552	%
The percentage of pixels in not cloud (-1) class is 39.78950	%
The percentage of pixels in unlabeled (0) class is 23.43499	%

```{r}
ggplot(image1b, aes(x, y, colour = expert_label)) + 
  geom_point()
```
## (c)










# 2 Preparation

# 3 Modeling

# 4 Diagnostics 



***
